Large Linear Multi-output Gaussian Process Learning for Time Series

### 2017-07-31 paper/paper.tex
Fixed corrections from Prof. Cunningham
1. Section 2.2, O(n^{3/P}) -> O(Pn^{(1+P)/P})
2. Section 4, emphasized how stationarity still plays a role in our work

If we can identify structure in the covariance $K$, then we can develop fast in-memory representations and efficient matrix-vector multiplications (MVMs) for $K$---this has been used in the past to accelerate GP model estimation \cite{gilboa2015scaling, cunningham2008fast}. The Kronecker product $A\otimes B$ of matrices of order $a,b$ is a block matrix of order $ab$ with $ij$th block $A_{ij}B$. We can represent the product by keeping representations of $A$ and $B$ separately, rather than the product. Then, the corresponding MVMs can be computed in time $O(a\mvm(B)+b\mvm(A))$, where $\text{MVM}(\cdot)$ is the runtime of a MVM. For GPs on uniform dimension-$P$ grids, this reduces the runtime of finding $\mcL$ from $O(n^3)$ to $O(n^{3/P})$ \cite{gilboa2015scaling}.

@@

With multiple outputs, non-differentiable cross-covariance interactions violate the assumptions of SKI's cubic interpolation. We show a modification to SKI is compatible with the piecewise-differentiable LMC kernel. This approximation induces the previously-identified block-Toeplitz structure of the LMC kernel, enabling acceleration of GP estimation for non-uniform inputs.

>>>>

If we can identify structure in the covariance $K$, then we can develop fast in-memory representations and efficient matrix-vector multiplications (MVMs) for $K$---this has been used in the past to accelerate GP model estimation \cite{gilboa2015scaling, cunningham2008fast}. The Kronecker product $A\otimes B$ of matrices of order $a,b$ is a block matrix of order $ab$ with $ij$th block $A_{ij}B$. We can represent the product by keeping representations of $A$ and $B$ separately, rather than the product. Then, the corresponding MVMs can be computed in time $O(a\mvm(B)+b\mvm(A))$, where $\text{MVM}(\cdot)$ is the runtime of a MVM. For GPs on uniform dimension-$P$ grids, this reduces the runtime of finding $\mcL$ from $O(n^3)$ to $O\pa{Pn^{\nicefrac{1+P}{P}}}$ \cite{gilboa2015scaling}.

@@

With multiple outputs, LMC cross-covariance interactions violate the assumptions of SKI's cubic interpolation, which require full stationarity (invariance to translation in the indexing space $\mcX'$). We show a modification to SKI is compatible with the piecewise-differentiable LMC kernel, which is only invariant to translation along the indexing subspace $\mcX$ (the subkernels $k_q$ are stationary). This partial stationarity induces the previously-identified block-Toeplitz structure of the LMC kernel and enables acceleration of GP estimation for non-uniform inputs.
